The Chemistry Development Kit (CDK). 3. Atom typing, Rendering, Molecular Formula, and Substructure Searching

### 2016-08-23 article.tex
Added Stefan Kuhn
Added the RDT reference (closes #26)

\author[
   addressref={um},                   % id's of addresses, e.g. {aff1,aff2}
   %corref={aff1},                       % id of corresponding address, if any
   %noteref={n1},                        % id's of article notes, if any
   email={egon.willighagen@maastrichtuniversity.nl}   % email address
]{\inits{ELW}\fnm{Egon L} \snm{Willighagen}}
\author[
   addressref={aff2},
   email={john@nextmovesoftware.com}
]{\inits{JWM}\fnm{John W} \snm{May}}
\author[
   addressref={uppsala},
   email={jonathan.alvarsson@farmbio.uu.se}
]{\inits{JA}\fnm{Jonathan} \snm{Alvarsson}}
\author[
   addressref={uppsala},
   email={berg.arvid@gmail.com}
]{\inits{AB}\fnm{Arvid} \snm{Berg}}
\author[
   addressref={jena},
   email={kai.duehrkop@uni-jena.de}
]{\inits{KD}\fnm{Kai} \snm{Dührkop}}
\author[
   addressref={idea},
   email={jeliazkova.nina@gmail.com}
]{\inits{NJ}\fnm{Nina} \snm{Jeliazkova}}
\author[
   addressref={wi_mit},
   email={pluskal@wi.mit.edu}
]{\inits{TP}\fnm{Tomáš}~\snm{Pluskal}}
\author[
   addressref={???},
   email={miguelrojasch@googlemail.com}
]{\inits{MRJ}\fnm{Miguel}~\snm{Rojas-Cherto}}
\author[
   addressref={uppsala},
   email={ola.spjuth@farmbio.uu.se}
]{\inits{OS}\fnm{Ola} \snm{Spjuth}}
\author[
   addressref={???},
   email={gilleain.torrance@gmail.com}
]{\inits{GT}\fnm{Gilleain} \snm{Torrance}}
\author[
   addressref={um},
   email={chris.evelo@maastrichtuniversity.nl}
]{\inits{CTE}\fnm{Chris T} \snm{Evelo}}
\author[
   addressref={nih},
   email={guhar@mail.nih.gov}
]{\inits{RG}\fnm{Rajarshi} \snm{Guha}}
\author[
   addressref={ebi},
   email={steinbeck@ebi.ac.uk}
]{\inits{CS}\fnm{Christoph}~\snm{Steinbeck}}

\address[id=um]{
  \orgname{Dept of Bioinformatics - BiGCaT, NUTRIM, Maastricht University}, % university, etc
  %\street{Waterloo Road},                     %
  \postcode{NL-6200 MD},
  \city{Maastricht},                              % city
  \cny{The Netherlands}                                    % country
}
\address[id=aff2]{
  \orgname{??},
  %\street{D\"{u}sternbrooker Weg 20},
  %\postcode{24105},
  %\city{Kiel},
  \cny{UK}
}
\address[id=uppsala]{
  \orgname{Department of Pharmaceutical Biosciences, Uppsala University},
  %\street{D\"{u}sternbrooker Weg 20},
  \postcode{751 24},
  \city{Uppsala},
  \cny{Sweden}
}
\address[id=jena]{
  \orgname{Chair for Bioinformatics, Friedrich Schiller University},
  %\street{},
  \postcode{07743},
  \city{Jena},
  \cny{Germany}
}
\address[id=idea]{
  \orgname{Ideaconsult Ltd},
  \street{A. Kanchev 4},
  \postcode{1000},
  \city{Sofia},
  \cny{Bulgaria}
}
\address[id=wi_mit]{
  \orgname{Whitehead Institute for Biomedical Research}, 
  \street{9 Cambridge Center},
  \postcode{MA 02142},
  \city{Cambridge},
  \cny{USA}
}
\address[id=nih]{
  \orgname{National Center for Advancing Translational Science},
  \street{9800 Medical Center Drive},
  \postcode{MD 20878},
  \city{Rockville},
  \cny{USA}
}
\address[id=ebi]{
  \orgname{Chemoinformatics and Metabolism team, European Bioinformatics Institute},
  %\street{A. Kanchev 4},
  %\postcode{1000},
  \city{Hinxton},
  \cny{UK}
}

@@

Open Source cheminformatics has made significant steps forward recently~\cite{OBoyle2011}.
The Chemistry Development Kit (CDK) is one of the tools under the Blue Obelisk wings,
and previously documented version have been widely adopted~\cite{Steinbeck2003,Steinbeck2006}.
For example, new tools expose CDK functionality in larger platforms such as Cinfony~\cite{OBoyle2008},
rcdk~\cite{Guha2007}, and plugins for Taverna~\cite{Truszkowski2011},
KNIME~\cite{Beisken2013}, ChemViz2 for Cytoscape~\cite{ChemViz2}, and LICSS for
Microsoft Excel~\cite{Lawson2012}.
Furthermore, projects with specific functionality have emerged building on the CDK, including
jCompoundMapper~\cite{Hinselmann2011}, ScaffoldHunter~\cite{wetzel2009interactive}, OMG~\cite{Peironcely2012},
Padel~\cite{yap2011padel}, AMBIT-SMARTS~\cite{jeliazkova2011ambitsmarts}, AMBIT-Tautomer~\cite{kochev2013ambit},
ReactPRED~\cite{ReactPRED}, SMSD~\cite{Rahman2009,Rahman2014}, WhichCyp~\cite{Rostkowski2013},
and MetaPrint2D~\cite{Carlsson2010}. Other tools used previous versions already, and follow
newer releases of the CDK, such as Bioclipse~\cite{spjuth2007bioclipse,
spjuth2009bioclipse} and AMBIT~\cite{jeliazkova2011ambit}. While the above
tools suggest that the CDK is primarily used to create general tools, it is
important to note that it is also used to solve specific questions, like finding
the maximally bridging rings in chemical structures~\cite{Marth2015}.

>>>>

\author[
   addressref={um},                   % id's of addresses, e.g. {aff1,aff2}
   %corref={aff1},                       % id of corresponding address, if any
   %noteref={n1},                        % id's of article notes, if any
   email={egon.willighagen@maastrichtuniversity.nl}   % email address
]{\inits{ELW}\fnm{Egon L} \snm{Willighagen}}
\author[
   addressref={aff2},
   email={john@nextmovesoftware.com}
]{\inits{JWM}\fnm{John W} \snm{May}}
\author[
   addressref={uppsala},
   email={jonathan.alvarsson@farmbio.uu.se}
]{\inits{JA}\fnm{Jonathan} \snm{Alvarsson}}
\author[
   addressref={uppsala},
   email={berg.arvid@gmail.com}
]{\inits{AB}\fnm{Arvid} \snm{Berg}}
\author[
   addressref={jena},
   email={kai.duehrkop@uni-jena.de}
]{\inits{KD}\fnm{Kai} \snm{Dührkop}}
\author[
   addressref={idea},
   email={jeliazkova.nina@gmail.com}
]{\inits{NJ}\fnm{Nina} \snm{Jeliazkova}}
\author[
   addressref={leicester},
   email={shk12@le.ac.uk}
]{\inits{SK}\fnm{Stefan} \snm{Kuhn}}
\author[
   addressref={wi_mit},
   email={pluskal@wi.mit.edu}
]{\inits{TP}\fnm{Tomáš}~\snm{Pluskal}}
\author[
   addressref={???},
   email={miguelrojasch@googlemail.com}
]{\inits{MRJ}\fnm{Miguel}~\snm{Rojas-Cherto}}
\author[
   addressref={uppsala},
   email={ola.spjuth@farmbio.uu.se}
]{\inits{OS}\fnm{Ola} \snm{Spjuth}}
\author[
   addressref={???},
   email={gilleain.torrance@gmail.com}
]{\inits{GT}\fnm{Gilleain} \snm{Torrance}}
\author[
   addressref={um},
   email={chris.evelo@maastrichtuniversity.nl}
]{\inits{CTE}\fnm{Chris T} \snm{Evelo}}
\author[
   addressref={nih},
   email={guhar@mail.nih.gov}
]{\inits{RG}\fnm{Rajarshi} \snm{Guha}}
\author[
   addressref={ebi},
   email={steinbeck@ebi.ac.uk}
]{\inits{CS}\fnm{Christoph}~\snm{Steinbeck}}

\address[id=um]{
  \orgname{Dept of Bioinformatics - BiGCaT, NUTRIM, Maastricht University}, % university, etc
  %\street{Waterloo Road},                     %
  \postcode{NL-6200 MD},
  \city{Maastricht},                              % city
  \cny{The Netherlands}                                    % country
}
\address[id=aff2]{
  \orgname{??},
  %\street{D\"{u}sternbrooker Weg 20},
  %\postcode{24105},
  %\city{Kiel},
  \cny{UK}
}
\address[id=uppsala]{
  \orgname{Department of Pharmaceutical Biosciences, Uppsala University},
  %\street{D\"{u}sternbrooker Weg 20},
  \postcode{751 24},
  \city{Uppsala},
  \cny{Sweden}
}
\address[id=jena]{
  \orgname{Chair for Bioinformatics, Friedrich Schiller University},
  %\street{},
  \postcode{07743},
  \city{Jena},
  \cny{Germany}
}
\address[id=idea]{
  \orgname{Ideaconsult Ltd},
  \street{A. Kanchev 4},
  \postcode{1000},
  \city{Sofia},
  \cny{Bulgaria}
}
\address[id=wi_mit]{
  \orgname{Whitehead Institute for Biomedical Research}, 
  \street{9 Cambridge Center},
  \postcode{MA 02142},
  \city{Cambridge},
  \cny{USA}
}
\address[id=nih]{
  \orgname{National Center for Advancing Translational Science},
  \street{9800 Medical Center Drive},
  \postcode{MD 20878},
  \city{Rockville},
  \cny{USA}
}
\address[id=ebi]{
  \orgname{Chemoinformatics and Metabolism team, European Bioinformatics Institute},
  %\street{A. Kanchev 4},
  %\postcode{1000},
  \city{Hinxton},
  \cny{UK}
}
\address[id=leicester]{
  \orgname{Department of Informatics, University of Leicester},
  %\street{},
  %\postcode{LE1 7RH},
  \city{Leicester},
  \cny{UK}
}

@@

Open Source cheminformatics has made significant steps forward recently~\cite{OBoyle2011}.
The Chemistry Development Kit (CDK) is one of the tools under the Blue Obelisk wings,
and previously documented version have been widely adopted~\cite{Steinbeck2003,Steinbeck2006}.
For example, new tools expose CDK functionality in larger platforms such as Cinfony~\cite{OBoyle2008},
rcdk~\cite{Guha2007}, and plugins for Taverna~\cite{Truszkowski2011},
KNIME~\cite{Beisken2013}, ChemViz2 for Cytoscape~\cite{ChemViz2}, and LICSS for
Microsoft Excel~\cite{Lawson2012}.
Furthermore, projects with specific functionality have emerged building on the CDK, including
jCompoundMapper~\cite{Hinselmann2011}, ScaffoldHunter~\cite{wetzel2009interactive}, OMG~\cite{Peironcely2012},
Padel~\cite{yap2011padel}, AMBIT-SMARTS~\cite{jeliazkova2011ambitsmarts}, AMBIT-Tautomer~\cite{kochev2013ambit},
ReactPRED~\cite{ReactPRED}, SMSD~\cite{Rahman2009,Rahman2014,Rahman2016}, WhichCyp~\cite{Rostkowski2013},
and MetaPrint2D~\cite{Carlsson2010}. Other tools used previous versions already, and follow
newer releases of the CDK, such as Bioclipse~\cite{spjuth2007bioclipse,
spjuth2009bioclipse} and AMBIT~\cite{jeliazkova2011ambit}. While the above
tools suggest that the CDK is primarily used to create general tools, it is
important to note that it is also used to solve specific questions, like finding
the maximally bridging rings in chemical structures~\cite{Marth2015}.
